### 2017-05-30 pkg/model4you/inst/JORS/man_model4you.Rnw
abstract and start intro
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1397 255f4249-2263-4417-b1ea-8498779e27a7

@@

<<setup, echo=FALSE>>=
knitr::opts_chunk$set(cache = TRUE, message = FALSE)
@

@@

\section*{Abstract}

\textcolor{blue}{A short (ca. 100 word) summary of the software being described: what problem the software addresses, how it was implemented and architected, where it is stored, and its reuse potential.}

@@

\textcolor{blue}{An overview of the software, how it was produced, and the research for which it has been used, including references to relevant research articles. A short comparison with software which implements similar functionality should be included in this section. }

@@

<<GBSG2_prep>>=
library("model4you")
library("survival")
library("ggplot2")
library("gridExtra")
data("GBSG2", package = "TH.data")

>>>>

%% commands

@@

<<setup, echo=FALSE>>=
library("knitr")
knitr::opts_chunk$set(cache = TRUE, message = FALSE)

knit_hooks$set(inline = function(x) { 
  if (is.numeric(x)) return(knitr:::format_sci(x, 'latex')) 
  highr:::hi_latex(x) 
}) 
@

@@

\section*{Abstract}
Typical models estimating treatment effecs assume that the treatment effect is the same for all 
individuals observed and 
for possible future individuals. Model-based recursive partitioning allows to relax this assumption.
With model-based trees one can compute treatment effects for different strata of individuals. The
strata are found in a data driven fashion and depend on characteristics of the individuals. 
Model-based random forests allow for a similarity estimation between individuals in terms of 
model estimates and through this we can estimate personalised models.
The R package \emph{model4you} (available on CRAN) implements these stratified and personalised models with a
focus on ease of use and interpretability so that clinicians and other users can take
the model they usually use for the estimation of the average treatment 
effect and with a few lines of code get a visualisation that is easy to understand and interpret.

%\textcolor{blue}{A short (ca. 100 word) summary of the software being described: what problem the software addresses, how it was implemented and architected, where it is stored, and its reuse potential.}

@@

%\textcolor{blue}{An overview of the software, how it was produced, and the research for which it has been used, including references to relevant research articles. A short comparison with software which implements similar functionality should be included in this section. }
Studies in various fields randomly assign individuals one of two groups with
different exposure and then measure a response. For example, in clinical trials
patients are assigned to one of two treatment groups where usually one treatment
group receives a new treatment or drug and the other treatment group receives
the standard of care or a placebo.

Treatment effect estimation is often done using simple models with the binary
treatment indicator as only covariate. In the example of a clinical trial the
treatment indicator would be $1$ if the patient receives the new treatment and
$0$ if the patient receives standard of care.
In R such a simple model can be estimated as follows:
<<eval=FALSE>>=
base_model <- model(response ~ treatment, data)
@
with \Sexpr{'response'} being the response measured, \Sexpr{'treatment'} being the
treatment indicator and \Sexpr{'data'} being the data set containing these variables.
The function \Sexpr{'model()'} can be replaced for example by \Sexpr{'lm()'} to estimate
a linear model, \Sexpr{'glm()'} to estimate a generalised linear model or 
\Sexpr{'survreg()'} to estimate a parametric survival model. 
These models estimate intercept and treatment effect for
all individuals in the \Sexpr{'data'} and allow for predicting the response 
of other individuals given they do or don't receive the treatment of interest.

If you do not believe that all individuals have the same intercept and treatment
effect the R package \emph{model4you} gives you two options to loosen up this 
strict model: 
\paragraph{(1) Model-based trees} identify subgroups where within the subgroups the model 
parameters are similar and between groups the model parameters are different. 
<<eval=FALSE>>=
strat_models <- pmtree(base_model)
@

\paragraph{(2) Personalised models} use model-based random forests to estimate similarity
of individuals in terms of model parameters. For each individual a personalised model can
be estimated based on a weighted set of the original data, where the similarity measure 
corresponds to the weight.
<<eval=FALSE>>=
pers_models <- pmodel(base_model)
@

@@

<<GBSG2_prep>>=
library("model4you")
library("survival")
data("GBSG2", package = "TH.data")
